Enforced Invariant Confluence

### 2020-08-09 doc/the_vldb_journal_paper/rebuttal.tex

  \item (Section 6.4)
    We added a paragraph explaining how a segmented invariant confluence
    decision procedure can leverage our merge reducibility results in addition
    to our results on invariant closure.

  \item (Section 6.4)
    We clarified the fault tolerance limitations of segmented invariant
    confluence. A naive implementation of our segmented invariant confluence
    decision procedure does not guarantee liveness, but there are
    straightforward techniques we can leverage to introduce liveness. We added
    an explanation of this to the paper.

  \item (Section 7)
    Our conference paper presents formal definitions of distributed objects and
    invariant confluence. The definitions revolve around servers gossiping
    state to one another and merging the states together using a binary merge
    operator. This approach is called a ``state-based'' approach. We added a
    section (Section 7) outlining a dual ``operation-based'' approach in which
    transactions, rather than states, are gossiped. This duality between
    state-based and operation-based definitions is well established in the CRDT
    literature.
\end{itemize}

\paragraph{Reviewer 1: D7}
\begin{feedback}
  In the intro we read about ``the trade-off between application complexity and
  coordination-freedom'', it would be great to see some practical examples later
  in the paper
\end{feedback}
The trade-off between application complexity and coordination-freedom is shown
in Examples 2, 6, 12, and 13. Very simple examples (Example 2) don't require
any segmentation. More complex initial states (Example 6), more complex
transactions (Example 12), and more complex optimizations (Example 13) require
segmentation. We agree though that these examples don't demonstrate the
trade-off in an airtight way. The trade-off we describe in the introduction is
written to give some intuition to the reader, but ultimately the notion of
complexity is subjective, and the number of segments of an application is up to
the user.

\paragraph{Reviewer 1: D8}
\begin{feedback}
  In section 4.2, more details about what can be done with respect to help a
  user reason about unreachable states would be great
\end{feedback}
See Reviewer 3: Section 4.2.

### 2020-08-10 doc/the_vldb_journal_paper/rebuttal.tex

% Turn on the full reviewer comments if the turnonshowfeedback flag was passed
% in from the command line. This was taken from [1].
%
% [1]: https://stackoverflow.com/a/1466610/3187068
% \ifdefined\turnonshowfeedback
%   \toggletrue{showfeedback}
% \else
%   \togglefalse{showfeedback}
% \fi

\begin{document}
\begin{center}
  \ttlfnt Response to Reviewer Comments
\end{center}

\begin{multicols*}{2}
Below, you'll find our response to reviewer comments and our revisions. All
changes made to the paper are shown in \markrevisions{pink}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Overview}
Our original publication on \invariantconfluence{} is largely theoretical. We
provide new, and arguably improved, definitions of \invariantconfluence{}. We
characterize the relationship between \invariantclosure{} and
\invariantconfluence{}. We exploit the relationship to develop an
\invariantclosure{} decision procedure. We invent the notion of
\mergereducibility{} and characterize its relationship with
\invariantclosure{}. We invent the notion of segmented \invariantconfluence{},
and develop another decision procedure.

When our paper was invited to this journal, we took the opportunity to improve
the rigor and clarity of the core contributions of the paper. We added numerous
proofs to verify the theorems in the paper, and we added many examples and
illustrations providing valuable intuition behind the key results. We also
added a significant amount of new material:
\begin{itemize}
  \item (Section 2.3)
    In our conference paper, we present a definition of invariant confluence
    that is slightly different than the original definition from Bailis et
    al.~\cite{bailis2014coordination}, but we do not comment on the difference.
    We added a subsection (Section 2.3) clarifying that the difference in
    definitions is superficial. We describe Bailis' original definition using
    our formalism and prove that the two definitions are equivalent.

The requested revisions, however, fall mainly into two categories. First, we
have requests to expand our empirical evaluation. Unfortunately, an expanded
evaluation requires a significant amount of work for very little additional
technical contribution. The applications of our theoretical results are
important (which is why we included them in the paper originally), but it is
not the focus of the paper. We believe that future work focused more on the
application of \invariantconfluence{} would leverage the key insights of our
paper and would benefit from the increased clarity and rigor of our journal
submission.

Second, we have requests to complete some of our suggested future work, with
arguments that the two years since our original publication should be enough
time to do so. We disagree that the time since our original publication should
have any bearing on our journal submission. Since our publication, we have been
working in new research areas. We have not spent the last two years working on
\invariantconfluence{}. Moreover, the suggested future work involves a
significant research undertaking that would involve solving many fundamental
research problems with implications well beyond \invariantconfluence{}. We
believe that any such results would more than warrant a publication on their
own.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 1}
\paragraph{Reviewer 1: W1}
\begin{feedback}
  More details about the DSL to specify the invariants with more complex
  examples would add a lot of value to the paper
\end{feedback}
We added a link to our DSL implementation on GitHub to the beginning of Section
8. We also added a bit more explanation of the DSL to Section 8.1.  The paper
already has over 10 examples of invariant confluence with varying degrees of
complexity from simple (e.g., Example 10) to complex (e.g., Example 14), and we
feel that additional examples will offer little additional technical insight.

\paragraph{Reviewer 1: W2}
\begin{feedback}
  More experiments with some realistic workloads (like TPCC) to show the
  benefit of the proposed concepts in action are needed
\end{feedback}
We agree that having a number of realistic workloads in the paper would be
beneficial, but specifying, verifying, and explaining every workload requires a
significant amount of work for a marginal increase in the clarity of the paper.
The focus of the paper is the foundations, not the applications of these ideas.

@@

\paragraph{Reviewer 1: D8}
\begin{feedback}
  In section 4.2, more details about what can be done with respect to help a
  user reason about unreachable states would be great
\end{feedback}
See Reviewer 3: Section 4.2.

\paragraph{Reviewer 1: D9}
\begin{feedback}
  Figure 5 should be moved before figure 6
\end{feedback}
Fixed!

\paragraph{Reviewer 1: D10}
\begin{feedback}
  in page 3 we read ``The culprit is concurrent decrements, which we can forbid
  in a simple one-segment segmentation ...'', this sentence does not parse well
\end{feedback}
We reworded the sentence to be easier to read.

\paragraph{Reviewer 1: D11}
\begin{feedback}
  In 6.4 we read ``This example is exactly the same as Example 8 which we
  handled by simply removing withdrawal transactions from our segmentation's
  set of transactions.'' How can one identify transactions to remove, and is
  the removal practically done?
\end{feedback}
In the current work, users must manually identify transactions to remove. More
generally, identifying a transaction to remove from a segmentation is a special
case of forming a segmentation, and we do not provide any automated assistance
on how to form a segmentation. Is the removal practically done? That depends on
the workload. For some workloads, identifying invariant-violating transactions
can be easy, like the bank withdrawal example (Example 8). Other, more complex
workloads, can be difficult to reason about. Forming segmentations for these
workloads is not easy. We think that automatically forming segmentations is an
interesting avenue for future research.

\paragraph{Reviewer 1: D12}
\begin{feedback}
  In 8.1 we read ``When a server receives a coordination request from the
  leader, it stops processing transactions and sends the leader its state.''
  what happens to the in-flight transactions? Are the cancelled or allowed to
  finish? Is the system simply not admitting new transactions?
\end{feedback}
Thanks for catching this. We added some clarification to Section 8.1.  In
short, in-flight transactions are deferred until after the round of
coordination is done. They aren't cancelled, and they are allowed to finish,
but only after the coordination is done. The entire system stops processing
transactions temporarily. Technically, the system can still \emph{admit} new
transactions during coordination, but it cannot execute them.

@@

Your question also reinforces the conclusion of Figure 10. If coordination is
frequent, then segmented \invariantconfluence{} becomes slow, even slower than
linearizable replication. Segmented \invariantconfluent{} works well only when
coordination is infrequent.

\paragraph{Reviewer 1: D13}
\begin{feedback}
  In 8.1 again, at the end, more discussion about the tradeoff of
  fault-tolerance would be interesting.
\end{feedback}
See Reviewer 4: 5, 7.

\paragraph{Reviewer 1: Typos}
\begin{feedback}
  end of first column of page 3 ``that it execute a transaction'' $\to$ ``that
  it executes a transaction''

  section 5, in example 3, ``Our start state state'' $\to$ two times state

  section 8.2 in example 12: ``and a optional'' $\to$ ``and an optional''
\end{feedback}
Fixed.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 3}
\paragraph{Reviewer 3: Overview}
\begin{feedback}
  This manuscript is an extension of the VLDB 2019 conference version of
  ``Interactive Checks for Coordination Avoidance''. The original PVLDB paper
  is a solid piece of work that is well written and easy to follow. It was
  published in September 2018. In this manuscript, the authors elaborated more
  by adding more examples, proofs, and figures to make it easier for the reader
  to understand specific concepts and definitions. Some readers may find that
  the original paper is clear and enough without these additions. Moreover, the
  authors added two new sections, namely Section 4.3 and 7, which do not show a
  significant improvement to the original contribution. In almost two years, it
  was expected that the authors will address some of the future work mentioned
  in the original paper, such as ``In the future, we plan on exploring ways to
  help a user reason about unreachable states and ways to discover sets of
  unreachable states automatically.'' Overall, the authors did not manage to
  show that the additions were significant enough to warrant publication.
\end{feedback}
We address this feedback in detail below. We also disagree that the time
between publication and now should have any bearing on the revisions.

\paragraph{Reviewer 3: Section 2.1}
\begin{feedback}
  The authors revised a small part of this section to clarify that the merge
  operator of a distributed object does not have to satisfy any special
  mathematical properties like associativity and commutativity.
\end{feedback}
Yup. This was requested by many readers who assumed early on that the merge
operator did require some special properties. In Bailis et al.'s original
paper, for example, they assume the merge operator is associative, commutative,
and idempotent~\cite{bailis2014coordination}.

\paragraph{Reviewer 3: Section 2.3}
\begin{feedback}
  The authors added this section to prove that their definition of invariant
  confluence and the original definition from Bailis et al.\ are equivalent.
  The authors did not clarify why it is important to prove that both
  definitions are equivalent.
\end{feedback}
We believe that proving equivalent definitions are, in fact, equivalent is a
fundamental tenant of theoretical research. Without a proof of the two
definitions being equivalent, it's possible that our notion of
\invariantconfluence{} and Bailis et al.'s original notion differ.  If so, none
of our theorems or decision procedures could be applied to Bailis et al.'s
findings. We'd be talking about two completely different concepts.  Proving
they are equivalent unifies the work. Without this knowledge, proofs using the
different definitions would be mutually incompatible.

\paragraph{Reviewer 3: Section 3}
\begin{feedback}
  The authors revised a small part of the original section to be in the form of
  a proof. For example in the PVLDB paper, the authors said: ``The forward
  direction of Theorem 2 follows immediately from Theorem 1. The backward
  direction holds because any two invariant satisfying states s1 and s2 must be
  reachable (by assumption), so their join s1 t s2 is also reachable. And
  because O is (s0, T, I)-confluent, all reachable points, including s1 t s2,
  satisfy the invariant.'' And in the VLDBJ submission, the authors said:
  ``Proof. The forward direction of Theorem 3 follows immediately from Theorem
  2. The backward direction holds because any two invariant satisfying states
  s1 and s2 must be reachable (by assumption), so their join s1 ⊔ s2 is also
  reachable. And because O is (s0, T, I)- confluent, all reachable points,
  including s1 ⊔s2, satisfy the invariant. Moreover s0 is reachable, so I(s0)
  because O is (s0,T,I)-confluent.''
\end{feedback}
Yup, a minor but (we hope) useful improvement.

@@

\paragraph{Reviewer 3: Section 4.1}
\begin{feedback}
  the authors expanded their correctness proof of the interactive decision
  procedure and added a few other minor elaborations. The original proof in the
  PVLDB paper was clear enough to the reader.
\end{feedback}
We're happy that the original proof was clear to the reviewer, but we felt it
was not clear to most people, including us.

\paragraph{Reviewer 3: Section 4.2}
\begin{feedback}
  In Section 4.2, the authors added a very brief example showing why invariant
  closure is undecidable. The reader can find original section 4.2 is clear
  enough. Instead in both versions of section 4.2, i.e., PVLDB and VLDBJ, the
  authors said: ``In the future, we plan on exploring ways to help a user
  reason about unreachable states and ways to discover sets of unreachable
  states automatically.'' After almost two years of publishing the PVLDB
  version of this submission, this manuscript would be much stronger, if the
  authors explored these ways.
\end{feedback}
Automatically discovering unreachable states is hard, really hard. As we
describe in the paper, it is easy to verify that a state is reachable, simply
give an expression that evaluates to it. Verifying that a state is unreachable
is significantly harder, as we have to create a proof showing that there does
not exist any execution in which the state is reached. Open mathematical
problems like the Collatz conjecture~\cite{mottaanalysis} show that reasoning
about the reachability or unreachability of even simple programs can be
tremendously difficult. If we developed any substantial techniques to identify
unreachable states in a general way, we feel that would constitute a strong
publication on its own.

@@

\paragraph{Reviewer 4: 4}
\begin{feedback}
  Section 4 is about invariant tightening, which has a broader applicability in
  distributed systems/programs. Unfortunately the techniques used here are not
  new. Involving the user for identifying unreachable states in not practical,
  and the discussion on tolerating user errors is weak as it does not offer a
  remedy. Would using some heuristics and abstracting state provide any help
  here?
\end{feedback}
We agree that involving the user for identifying unreachable states is not
ideal, and in many cases not practical. We think our discussion of tolerating
user errors is also not ideal, but not weak. By requiring users to provide
proofs of reachability and by deploying the invariant $I - NR$ instead of $I$,
invariant violations are impossible.
%
Heuristics and abstracting state and other techniques from invariant tightening
would probably help, but they are out of scope of the paper. As you mention,
the techniques are not new. We view it as an orthogonal but complementary line
of research.

\paragraph{Reviewer 4: 5}
\begin{feedback}
  Section 6 on segmented invariant confluence is nice. It would be nice if
  Section 6.4 could be improved. I liked the fourth point in Section 6.4, would
  it be possible to elaborate on this. The sixth point however is bad, as
  replicating every server with state-machine-replication is overkill. Would it
  be possible to use use a membership service approach (as in Vertical Paxos
  and Chain Replication configuration) to prevent a single crash to flunk
  global coordination?
\end{feedback}
We added a couple paragraphs to Example 13 with a concrete example of the
optimization described in our fourth point. We also added a forward reference
in Section 6.4 to Example 13.

Your comment on our sixth point is interesting. We hadn't thought of that
before. Coincidentally, since our initial publication, we spent some time
researching membership changes in state machine replication protocols---which
you can find here:
\url{https://mwhittaker.github.io/publications/matchmaker_paxos.html}---so we
can answer your question with quite a bit of confidence. Assume a client sends
a transaction to a server $p$. The transaction happens to not need global
coordination, so $p$ executes the transaction locally and sends back the result
to the client. This scenario makes membership changes difficult for the
following reason. Consider a different execution in which $p$ is partitioned
from the rest of the servers. The system erroneously thinks that $p$ has failed
and reconfigures to some other set of servers. However, $p$ and the client are
oblivious to this. The client can still contact $p$ and have transactions
executed, even though the system has moved to a completely new set of servers.
This can lead to situations that are impossible without membership changes. For
example, $p$ may be in a different active segment than the rest of the servers.

### 2020-08-10 doc/the_vldb_journal_paper/sections/evaluation.tex
Hopefully final revisions...
...before catching typos of course :)

Lucy also includes an implementation of the \invariantconfluence{} and
segmented \invariantconfluence{} system models in roughly 3,500 lines of C++.
Users specify objects, transactions, invariants, and segmentations in C++. Lucy
then replicates the objects using segmented \invariantconfluence{}. Clients
send every transaction request to a randomly selected server. When a server
receives a transaction request, it executes \algoref{TxnExecution} to attempt
to execute the transaction locally.  If the transaction requires global
coordination, then the server forwards the transaction request to a
predetermined leader. When the leader receives a transaction request, it
broadcasts a coordination request to the other servers. When a server receives
a coordination request from the leader, it stops processing transactions and
sends the leader its state. When the leader receives the states of all other
servers, it executes the transaction, and then sends its state to the other
servers. When a server receives a new state, it adopts the state, computes its
new active segment, and resumes normal processing. After every 100 transactions
processed, a server sends a merge request to a randomly selected server.

>>>>

Lucy also includes an implementation of the \invariantconfluence{} and
segmented \invariantconfluence{} system models in roughly 3,500 lines of C++.
Users specify objects, transactions, invariants, and segmentations in C++. Lucy
then replicates the objects using segmented \invariantconfluence{}. Clients
send every transaction request to a randomly selected server. When a server
receives a transaction request, it executes \algoref{TxnExecution} to attempt
to execute the transaction locally.  If the transaction requires global
coordination, then the server forwards the transaction request to a
predetermined leader. When the leader receives a transaction request, it
broadcasts a coordination request to the other servers. When a server receives
a coordination request from the leader, it stops processing transactions and
sends the leader its state. \markrevisions{All in-flight transactions are
deferred until the global communication is complete.} When the leader receives
the states of all other servers, it executes the transaction, and then sends
its state to the other servers. When a server receives a new state, it adopts
the state, computes its new active segment, and resumes normal processing.
After every 100 transactions processed, a server sends a merge request to a
randomly selected server.
